---
title: "leapR"
output: html_notebook
---

# LeapR


### Figure 2
We will compare the ability of transcriptomics, proteomics, and phosphoproteomics to inform about
differences between short and long surviving patient groups.

This spans the multiple enrichment methods in leapR and also includes multi-omics

The resulting heatmap is presented as Figure 2 in the paper.

```{r}
proteomics <- read.csv("bonita_proteomics.csv", row.names = 1, check.names = 1)
phosphoproteomics <- read.csv("bonita_phosphoproteomics.csv", row.names = 1, check.names = 0)
transcriptomics <- read.csv("bonita_transcriptomics.csv", row.names = 1, check.names =1)
contrasts <- read.csv("concatenated_conditions_CAMERA.csv", row.names = 1, check.names = FALSE)
prot_contrast <- contrasts[colnames(proteomics),]
mrna_contrast <- contrasts[colnames(transcriptomics),]
phosph_contrast <- contrasts[colnames(phosphoproteomics),]
```
# load the single omic and multi-omic pathway databases

```{r}

require(leapR)

#data("krbpaths")
#data("mo_krbpaths")
#data('msigdb')
krbpaths <- leapR::read_gene_sets("kegg_networks_edited.gmt")#("c2.cp.kegg.v7.5.1.symbols.gmt")
krbpaths$names <- str_to_title(trimws(gsub("\\s\\-\\sHomo\\sSapiens\\s\\(Human\\)", "", krbpaths$names)))
mo_krbpaths <- leapR::read_gene_sets("kegg_networks_edited.gmt")#("c2.cp.kegg.v7.5.1.symbols.gmt")
mo_krbpaths$names <- str_to_title(trimws(gsub("\\s\\-\\sHomo\\sSapiens\\s\\(Human\\)", "", mo_krbpaths$names)))
#msigdb$names <- msigdb$names[grepl("KEGG", msigdb$names)]
#msigdb$desc <- msigdb$desc[grepl("KEGG", msigdb$names)]
#msigdb$sizes <- msigdb$sizes[grepl("KEGG", msigdb$names)]
#msigdb$matrix <- msigdb$matrix[msigdb$names[grepl("KEGG", msigdb$names)]]
#mo_krbpaths$names <- mo_krbpaths$names[grepl("KEGG", mo_krbpaths$names)]
#mo_krbpaths$desc <- mo_krbpaths$desc[grepl("KEGG", mo_krbpaths$names)]
#mo_krbpaths$sizes <- mo_krbpaths$sizes[grepl("KEGG", mo_krbpaths$names)]
#mo_krbpaths$matrix <- mo_krbpaths$matrix[mo_krbpaths$names[grepl("KEGG", mo_krbpaths$names)]]
```

# Run leapR analysis

```{r}
run_leapR <- function(shortlist, longlist, contrast1, contrast2) {
  # comparison enrichment in transcriptional data
  transdata.comp.enrichment.svl = leapR(
    geneset = krbpaths,
    enrichment_method = 'enrichment_comparison',
    datamatrix = transcriptomics,
    primary_columns = intersect(colnames(transcriptomics), shortlist),
    secondary_columns = intersect(colnames(transcriptomics), longlist), min_p_threshold=1
  )
  transdata.comp.enrichment.svl$Method <- "Transcriptomics"
  
  
  # comparison enrichment in proteomics data
  # this is the same code used above, just repeated here for clarity
  protdata.comp.enrichment.svl = leapR(
    geneset = krbpaths,
    enrichment_method = 'enrichment_comparison',
    datamatrix = proteomics,
    primary_columns = intersect(colnames(proteomics), shortlist),
    secondary_columns = intersect(colnames(proteomics), longlist), min_p_threshold=1
  )
  protdata.comp.enrichment.svl$Method <- "Proteomics"
  
  # comparison enrichment in phosphoproteomics data
  phosphodata.comp.enrichment.svl = leapR(
    geneset = krbpaths,
    enrichment_method = 'enrichment_comparison',
    datamatrix = phosphoproteomics,
    primary_columns = intersect(colnames(phosphoproteomics), shortlist),
    secondary_columns = intersect(colnames(phosphoproteomics), longlist), min_p_threshold=1
  )
  phosphodata.comp.enrichment.svl$Method <- "Phosphoproteomics"
  
  prot2 <- proteomics
  colnames(prot2) <-
    unlist(lapply(colnames(prot2), function(x) {
      return(names(as.list(contrasts[x, ])[as.list(contrasts[x, ]) == 1]))
    }))
  mrna2 <- transcriptomics
  colnames(mrna2) <-
    unlist(lapply(colnames(mrna2), function(x) {
      return(names(as.list(contrasts[x, ])[as.list(contrasts[x, ]) == 1]))
    }))
  
  phosph2 <- phosphoproteomics
  colnames(phosph2) <-
    unlist(lapply(colnames(phosph2), function(x) {
      return(names(as.list(contrasts[x, ])[as.list(contrasts[x, ]) == 1]))
    }))
  
  
  combodata <-
    combine_omics(
      proteomics = prot2,
      transcriptomics = mrna2,
      phospho = phosph2,
      id_column = 1
    )
  
  # comparison enrichment for combodata
  combodata.enrichment.svl = leapR(
    geneset = mo_krbpaths,
    enrichment_method = 'enrichment_comparison',
    datamatrix = combodata,
    primary_columns = c(contrast1),
    secondary_columns = c(contrast2),
    id_column = 1, min_p_threshold=1
  )
  combodata.enrichment.svl$pvalue[is.na(combodata.enrichment.svl$pvalue)] <-
    1
  #combodata.enrichment.svl <- combodata.enrichment.svl[combodata.enrichment.svl$pvalue < 1,]
  head(combodata.enrichment.svl)
  min(combodata.enrichment.svl$pvalue)
  max(combodata.enrichment.svl$pvalue)

  combodata.enrichment.svl$Method <- "Combined"
  head(combodata.enrichment.svl)
  min(combodata.enrichment.svl$pvalue)
  max(combodata.enrichment.svl$pvalue)
  
  transdata.comp.enrichment.svl <- transdata.comp.enrichment.svl[grepl("KEGG", rownames(transdata.comp.enrichment.svl)), ]
  temp <- !is.na(transdata.comp.enrichment.svl$BH_pvalue) #PValue < 0.01
  transdata.comp.enrichment.svl <- transdata.comp.enrichment.svl[temp,]
  transdata.comp.enrichment.svl$BH_pvalue <- p.adjust(transdata.comp.enrichment.svl$pvalue, "BH")
  
  protdata.comp.enrichment.svl <- protdata.comp.enrichment.svl[grepl("KEGG", rownames(protdata.comp.enrichment.svl)), ]
  temp <- !is.na(protdata.comp.enrichment.svl$BH_pvalue) #PValue < 0.01
  protdata.comp.enrichment.svl <- protdata.comp.enrichment.svl[temp,]
  
  protdata.comp.enrichment.svl$BH_pvalue <- p.adjust(protdata.comp.enrichment.svl$pvalue, "BH")
  
  
  phosphodata.comp.enrichment.svl <- phosphodata.comp.enrichment.svl[grepl("KEGG", rownames(phosphodata.comp.enrichment.svl)), ]
  temp <- !is.na(phosphodata.comp.enrichment.svl$BH_pvalue) #PValue < 0.01
  phosphodata.comp.enrichment.svl <- phosphodata.comp.enrichment.svl[temp,]
  
  phosphodata.comp.enrichment.svl$BH_pvalue <- p.adjust(phosphodata.comp.enrichment.svl$pvalue, "BH")
  combodata.enrichment.svl <- combodata.enrichment.svl[grepl("KEGG", rownames(combodata.enrichment.svl)), ]
  temp <- !is.na(combodata.enrichment.svl$BH_pvalue) #PValue < 0.01
  combodata.enrichment.svl <- combodata.enrichment.svl[temp,]
  
  combodata.enrichment.svl$BH_pvalue <- p.adjust(combodata.enrichment.svl$pvalue, "BH")
  
  transdata.comp.enrichment.svl$Pathway <- rownames(transdata.comp.enrichment.svl)
  protdata.comp.enrichment.svl$Pathway <- rownames(protdata.comp.enrichment.svl)
  phosphodata.comp.enrichment.svl$Pathway <- rownames(phosphodata.comp.enrichment.svl)
  combodata.enrichment.svl$Pathway <- rownames(combodata.enrichment.svl)
  
  results1 <-
    rbind(
      transdata.comp.enrichment.svl,
      protdata.comp.enrichment.svl,
      phosphodata.comp.enrichment.svl,
      combodata.enrichment.svl
    )
  return (results1)
}
```

# Contrast: "X19O2NoCyA - X1O2NoCyA"

``` {r, echo=TRUE, warning=FALSE, message=FALSE}
shortlist <- rownames(contrasts)[contrasts$`19O2NoCyA` == 1]
longlist <- rownames(contrasts)[contrasts$`1O2NoCyA` == 1]
contrast1 <- '19O2NoCyA'
contrast2 <- '1O2NoCyA'
results1 <- run_leapR(shortlist, longlist, contrast1, contrast2)
results1$Contrast <- "X19O2NoCyA - X1O2NoCyA"

head(results1)
```
# Contrast: "X1O2PlusCyA - X19O2NoCyA"

```{r}
shortlist <- rownames(contrasts)[contrasts$`1O2PlusCyA` == 1]
longlist <- rownames(contrasts)[contrasts$`1O2NoCyA` == 1]
contrast1 <- '1O2PlusCyA'
contrast2 <- '19O2NoCyA'
results2 <- run_leapR(shortlist, longlist, contrast1, contrast2)
results2$Contrast <- "X1O2PlusCyA - X19O2NoCyA"
head(results2)
```
# Contrast: "X1O2NoCyA - X1O2PlusCyA"

```{r}
shortlist <- rownames(contrasts)[contrasts$`1O2NoCyA` == 1]
longlist <- rownames(contrasts)[contrasts$`1O2PlusCyA` == 1]
contrast1 <- '1O2NoCyA'
contrast2 <- '19O2PlusCyA'
results3 <- run_leapR(shortlist, longlist, contrast1, contrast2)
results3$Contrast <- "X1O2NoCyA - X1O2PlusCyA"
head(results3)
```


```{r}
results <- rbind(results1, results2, results3)


results$negativelog10pvalue <- (-1)*log10(results$pvalue)
results$negativelog10FDR <- (-1)*log10(results$BH_pvalue)
temp <- !is.na(results$BH_pvalue) #PValue < 0.01
results <- results[temp,]
write.csv(results, "leapR_allResults.csv")

print(c(min(results$BH_pvalue[results$Method == "Transcriptomics"]), max(results$BH_pvalue[results$Method == "Transcriptomics"])))
print(c(min(results$BH_pvalue[results$Method == "Proteomics"]), max(results$BH_pvalue[results$Method == "Proteomics"])))


#temp <- results$BH_pvalue < 0.05 #PValue < 0.01
temp <- (results$pvalue < 0.05 & results$Method %in% c("Proteomics", "Transcriptomics")) | (results$BH_pvalue < 0.05 &results$Method %in% c("Phosphoproteomics"))
results <- results[temp,]
results$Pathway <- gsub("KEGG|_|\\d+$", " ", results$Pathway)
head(results)
dim(results)
```
# Plot results

```{r}
require(ggplot2)
ggplot(results) + geom_point(aes(x=negativelog10pvalue, y=Pathway, color=Contrast)) + facet_wrap(facets = 'Method', nrow = 1) + theme_bw() + xlab("-log10(UNCORRECTED P-VALUE)")
ggsave(filename = "leapR_allResults.svg", width = 12, height = 10, units = 'in')
ggsave(filename = "leapR_allResults.png", width = 12, height = 10, units = 'in')
```

